|

### 2020-03-10 paper/paper.Rmd
after making progress before deleting stuff
simulation went well and i think i have my final graph

```{r setup, include = FALSE}
# 
# https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html

@@

library(tidyverse)
library(ggtext)
library(here)
library(mirt)
library(cluster)
library(DescTools)
library(knitr)
library(kableExtra)
set.seed(123)
R.utils::sourceDirectory(here("R"))

@@

# Introduction

>>>>

```{r setup, include = FALSE}
# https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html

@@

library(tidyverse)
library(ggtext)
library(here)
library(mirt)
library(cluster)
library(DescTools)
library(knitr)
library(kableExtra)
set.seed(12345)
R.utils::sourceDirectory(here("R"))

@@

# Introduction {#intro}

### 2020-03-10 paper/paper_3_sim.Rmd
before deleting stuff
simulation went well and i think i have my final graph

@@

# i have no idea why this is like this...
out %>% 
  mutate(bigsim = bigsim %>% map(~ select(., foc_mean))) %>% 
  unnest(bigsim) %>% 
  ggplot(aes(x = n_dif_items, y = foc_mean)) +
  geom_point()

out$bigsim[[4]]$p[[1]] %>% colMeans()

```
Figure \ref{fig:difmap} provides intuition about the data generating process by showing the relationship between $\theta_i$ and $Pr(y_{ij} = 1)$ with $\eta_i$ set to the group mean. The items are ordered by the amount of DIF such that $\angle_{j = 11} = 30^\circ$ up to $\angle_{j = 20} = 60^\circ$.
```{r difmap, fig.cap = 'For a 20 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.', out.width="70%", warning = FALSE, message = FALSE}
out$bigsim[[2]]$pars[[4]] %>% 
  graph_pars_noncompensatory(0, -1) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = latex2exp::TeX("$\\theta_i$"),
    y = latex2exp::TeX("$Pr(y_{ij} = 1)$")
  )
```

Figure \ref{fig:simemmlg} shows the EM-MILG for one run using the same item parameters that generated Figure \ref{fig:difmap}. As expected, $\tilde{d_j}$ is about $\mu_\theta^\text{foc} - \mu_\theta^\text{ref} = -0.5 - 0 = -0.5$ for the first ten items which are DIF free. For the last ten items, $\tilde{d_j}$ increases as $\angle_j$ increases. Of course, we cannot include the "analyst chooses" method in our simulation, but regardless, the EM-MILG is a useful way to visualize the data.
```{r simemmlg, fig.cap = 'For a 20 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.', out.width="70%", warning = FALSE, message = FALSE}
out$bigsim[[2]]$intuitive_mod[[4]] %>%
    mod_intuitive_to_draws_df() %>%
    draws_df_to_logit_plot() +
    labs(
          x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$")
      )
```

@@

```{r achievegap, fig.cap = 'Each method's estimation of the target ability achievement gap.', out.width="70%", warning = FALSE, message = FALSE}
final_model_to_anchor_summary <- function(final_model, n_dif_items){
    coef_1f(final_model)$items %>%
        mutate(
            true_anchor =
                c(
                    rep("no dif", n_items - n_dif_items),
                    rep("some dif", ceiling(n_dif_items / 2)),
                    rep("high dif", floor(n_dif_items / 2))
                )
        ) %>%
        group_by(true_anchor) %>%
        summarize(selected = sum(anchor), total = n())
}

getfull <- function(one){
    one %>%
        select(ends_with("mod"), -intuitive_mod) %>%
        mutate(run = row_number()) %>%
        gather(method, model, -run) %>%
        mutate(
            n_anchors = model %>% map_int(~ sum(coef_1f(.)$items$anchor)),
            foc_mean = model %>% map_dbl(~ coef_1f(.)$ability$b_foc[1])
        ) %>%
        filter(n_anchors != 0 | method %in% c("gini_mod", "minbc_mod"))
}

out %>%
    mutate(achievegap = bigsim %>% map(getfull)) %>%
    select(-bigsim) %>%
    unnest(achievegap) %>%
    ggplot(aes(x = n_dif_items, y = foc_mean)) +
    geom_point(alpha = 0.5) +
    facet_wrap(~ method) +
    geom_hline(yintercept = -0.5, linetype = "dashed") +
    labs(
      x = "Number of DIF items",
      y = latex2exp::TeX("$\\hat{\\mu}_\\theta^foc$")
    )
```

>>>>

sim_n_dif_items_to_foc_mean <- function(sim, n_dif_items){
    mod <-
        mod_flexible(
            sim$data,
            sim$groups,
            flex_items = (ncol(sim$data) - n_dif_items + 1):ncol(sim$data)
        )

    coef_1f(mod)$ability$b_foc[1]
}

add_foc_mean <- function(one, n_dif_items){
    one %>%
    mutate(foc_mean = sim %>% map_dbl(sim_n_dif_items_to_foc_mean, n_dif_items))
}

@@

# i have no idea why this is like this...
out %>% 
  mutate(bigsim = bigsim %>% map(~ select(., foc_mean))) %>% 
  unnest(bigsim) %>% 
  ggplot(aes(x = n_dif_items, y = foc_mean)) +
  geom_point()
```
Figure \ref{fig:difmap} provides intuition about the data generating process by showing the relationship between $\theta_i$ and $Pr(y_{ij} = 1)$ with $\eta_i$ set to the group mean. The items are ordered by the amount of DIF such that $\angle_{j = 11} = 30^\circ$ up to $\angle_{j = 20} = 60^\circ$.
```{r difmap, fig.cap = 'For a 20 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.', out.width="70%", warning = FALSE, message = FALSE}
out$bigsim[[2]]$pars[[1]] %>% 
  graph_pars_noncompensatory(0, -1) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = latex2exp::TeX("$\\theta_i$"),
    y = latex2exp::TeX("$Pr(y_{ij} = 1)$")
  )
```

Figure \ref{fig:simemmlg} shows the EM-MILG for one run using the same item parameters that generated Figure \ref{fig:difmap}. As expected, $\tilde{d_j}$ is about $\mu_\theta^\text{foc} - \mu_\theta^\text{ref} = -0.5 - 0 = -0.5$ for the first ten items which are DIF free. For the last ten items, $\tilde{d_j}$ increases as $\angle_j$ increases. Of course, we cannot include the "analyst chooses" method in our simulation, but regardless, the EM-MILG is a useful way to visualize the data.
```{r simemmlg, fig.cap = 'For a 20 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.', out.width="70%", warning = FALSE, message = FALSE}
out$bigsim[[3]]$intuitive_mod[[1]] %>%
    mod_intuitive_to_draws_df() %>%
    draws_df_to_logit_plot() +
    labs(
          x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$")
      )
```

@@

```{r achievegap, fig.cap = 'Each method's estimation of the target ability achievement gap.', out.width="70%", warning = FALSE, message = FALSE}

n_items <- 12

final_model_to_anchor_summary <- function(final_model, n_dif_items){
    coef_1f(final_model)$items %>%
        mutate(
            true_anchor =
                c(
                    rep("no dif", n_items - n_dif_items),
                    rep("some dif", ceiling(n_dif_items / 2)),
                    rep("high dif", floor(n_dif_items / 2))
                )
        ) %>%
        group_by(true_anchor) %>%
        summarize(selected = sum(anchor), total = n())
}

getfull <- function(one){
    one %>%
        select(foc_mean, ends_with("mod"), -intuitive_mod) %>%
        mutate(run = row_number()) %>%
        gather(method, model, -run, -foc_mean) %>%
        mutate(
            n_anchors = model %>% map_int(~ sum(coef_1f(.)$items$anchor)),
            est_foc_mean = model %>% map_dbl(~ coef_1f(.)$ability$b_foc[1])
        ) %>%
        filter(n_anchors != 0 | method %in% c("gini_mod", "minbc_mod")) %>% 
    mutate(resid = est_foc_mean - foc_mean)
}

what <- 
  out %>%
    mutate(achievegap = bigsim %>% map(getfull)) %>%
    select(-bigsim) %>%
    unnest(achievegap)

what %>% 
  ggplot(aes(x = resid, y = as.factor(n_dif_items))) +
  ggridges::geom_density_ridges() +
  facet_wrap(~ method) +
  coord_flip() +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  labs(
      x = "Number of DIF items",
      y = latex2exp::TeX("$\\hat{\\mu}_\\theta^foc$")
  )
```
